---
title: "An overview of ArctosR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An overview of ArctosR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The ArctosR package can be used to download and save data from Arctos. This
vignette will provide an overview of the basic usage of this package.

```{r setup}
library(ArctosR)
```

## Basics

### Types of data in Arctos

Arctos contains records, such as specimens, and data associated with those
records.

### Queries and responses in ArctosR

ArctosR abstracts interacting with the Arctos API with objects. These objects
can be manipulated with the set of frontend functions for most tasks, or
created and manipulated directly by the user with builders.

#### Query objects

A **query** in ArctosR is a collection of searches by a user of the package for
a specific task. This might be: a simple search, such as a one-off search for
specimens of a species held by some museum; a search that requires the concatenation
of multiple responses, such as one that requests more records than Arctos can
provide in a single response; or a complex search that uses requests for one
set of search terms, then uses those responses as part of another request.

#### Response objects

A **response** then is an object that stores the contents of *one* response from
Arctos back to ArctosR. Each response has associated metadata, such as search
terms and timestamp, and content, such as a table of records matching those search
terms.

#### Metadata

Detailed **metadata** about each response is also saved in the user's query for
the purposes of documentation and reproducibility. This metadata is saved along
with downloaded records as a JSON file.

## Making queries to Arctos

### A basic query

ArctosR provides a single function, `get_records`, to search for records in the
Arctos database. In order to start building a search though, we have to find
out the possible *query parameters* we can use to search. These can be found with
the `get_query_parameters` function, which returns a dataframe of all
query parameters used by Arctos. The names in the `obj_name` column are what are
used as parameters to the `get_records` function.

```{r eval = FALSE}
> query_params <- get_query_parameters()
> query_params[1:6,1:3]
#>             display                obj_name category
#> 1     Verbatim Date           verbatim_date    event
#> 2 Collecting Method       collecting_method    event
#> 3 Collecting Source       collecting_source    event
#> 4        Ended Date              ended_date    event
#> 5  Event Attributes evtAttributeSearchTable    event
#> 6           Habitat                 habitat    event
```

For this basic query, we can use `guid_prefix`, whose description can be found
by listing the row of the query parameter dataframe:

```{r eval = FALSE}
> query_params[37,1:5]
#>       display    obj_name   category subcategory
#> 37 Collection guid_prefix identifier       basic
#>                                                                description
#> 37 Collection responsible for the record. Turning this off will break most
#>    forms.
```

We will also use `genus` and `species` whose description can be found at:

```{r eval = FALSE}
> query_params[28,1:5]
#>    display obj_name       category subcategory
#> 28   Genus    genus identification  curatorial 
#>                                               description
#> 28 Genus as provided in collection's preferred Source(s).

> query_params[23,1:5]
#>    display obj_name       category subcategory 
#> 23 Species  species identification  curatorial
#>                                                            description
#> 23 Species (binomial) as provided in collection's preferred Source(s).
```

Now that we have a set of parameters to use, we can pass them to the `get_records`
function and send our request to Arctos. This will return a **query**, which
bundles our search parameters with the returned data from Arctos.

```{r eval = FALSE}
query <- get_records(
  guid_prefix = "MSB:Mamm",
  genus = "Canis",
  species = "lupus")
```

### Requesting other columns than the core set

By default, the columns returned by Arctos are all of those with the category
core.

```{r eval = FALSE}
result_params <- get_result_parameters()
result_params[result_params$category == 'core',1:2]
#>                      display                      obj_name
#> 1  GUID (DarwinCore Triplet)                          guid
#> 7              Identified As               scientific_name
#> 43          Asserted Country                       country
#> 44   Asserted State/Province                    state_prov
#> 51         Specific Locality                 spec_locality
#> 57             Verbatim Date                 verbatim_date
#> 69          Decimal Latitude                       dec_lat
#> 70         Decimal Longitude                      dec_long
#> 71      Coordinate Error (m) coordinateuncertaintyinmeters
```

Additional columns can be requested by passing a vector of result parameters to
the `get_records` function in the `columns` parameter like so:

```{r eval = FALSE}
query <- get_records(guid_prefix = "MSB:Mamm", genus = "Canis", species = "lupus",
  columns = list(
    "guid", "scientific_name", "relatedcatalogeditems", "collectors",
    "state_prov", "spec_locality", "dec_lat", "dec_long", "verbatim_date",
    "examined_for", "detected", "not_detected"))
```

#### Requesting columns that are tables

Certain result parameters in Arctos are themselves tables, such as `partdetail`,
which links to the attributes of each part listed in `parts` associated with
a specimen. These can be requests just like any other result parameter, and
expanded into dataframes of their own by the frontend function `expand_column`.

```{r eval = FALSE}
query <- get_records(
  guid_prefix = "MSB:Mamm", genus = "Canis", species = "lupus",
  columns = list("guid", "parts", "partdetail"))
```

### Requesting all records

By default, `get_records` avoids requesting all records for a query unless
otherwise asked. By passing the parameter `all_records = TRUE` to `get_records`,
the user can request that ArctosR make multiple requests until all records for
a given query are downloaded.

```{r eval = FALSE}
get_record_count(
  guid_prefix = "MSB:Mamm",
  genus = "Canis",
  species = "lupus")
#> [1] 1694

query <- get_records(
  guid_prefix = "MSB:Mamm",
  genus = "Canis",
  species = "lupus",
  all_records = TRUE)
```

## Using data from Arctos

### Viewing data

```{r eval = FALSE}
query <- get_records(guid_prefix = "MSB:Mamm", genus = "Canis", species = "lupus",
  columns = list(
    "guid", "scientific_name", "parts", "collectors", "state_prov", "spec_locality",
    "dec_lat", "dec_long", "verbatim_date", "parts", "partdetail"))

msb_wolves <- response_data(query)
View(msb_wolves)
```

#### Expanding columns

```{r eval = FALSE}
expand_column(msb_wolves, "partdetail")
View(msb_wolves)
```

### Saving data

#### Saving as a flat CSV file

```{r eval = FALSE}
save_response_csv(msb_wolves, "msb_wolves.csv")
```

#### Saving an expanded CSV

```{r eval = FALSE}
save_response_csv(msb_wolves, "msb_wolves.csv", expanded = TRUE)
```

#### Saving as an RDS

```{r eval = FALSE}
save_response_rds(msb_wolves, "msb_wolves.rds")
```

#### Saving query metadata
